---
title: "RC_Statistical Analysis.Rmd"
author: "Matthew Ng"
date: "2023-09-21"
output:
  html_document:
    css:
    theme:       
      bg: "#1D1F20"
      fg: "#FFF"
      primary: "#007360"
      secondary: "#007360"
      success: "#007360"
    df_print: paged
    toc: true
    toc_float: true
    code_folding: show
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE}
library(knitr)
library(kableExtra)
thematic::thematic_rmd()
```

# Introduction
Hi! This is a R Cheatsheet done by Matthew HS Ng for everyone (and myself) as a guide for everything R that I have learnt thus far and for beginners to advanced users. Hope you find this helpful!

# Administrative Steps
We will first start by executing our administrative steps (loading packages, setting working directories, and importing our data). For more information on how to do that, please check out the Data Wrangling section of the R Cheatsheet Project

```{r}
#loading packages
library(dplyr) #data wrangling
library(psych) #descriptive statistics
library(lme4) #multilevel modelling
library(lmerTest) #multilevel modelling
library(corrtable) #for correlation table
library(car) #for additional applied regression support
```

```{r}
#loading datasets
setwd("C:/Users/Matthew/Desktop/R Cheatsheet Project")
dataset1 <- readxl::read_xlsx("RC_AnalysisPracticeDataset_1.xlsx")
#dataset2 <- 

```

Dataset 1 is a standard wide dataset with 500 observation and 18 variables. Variables include demograhpic information, self-report data and some physiological measures. This dataset is ideal for basic statistical analyses

Dataset 2: (TBC)

# Descriptive Statistics

For descriptive statistics, we usually use the psych package to extract the common descriptive statistics for the dataset
```{r}
#descriptive stats
dataset1_descriptives = psych::describe(dataset1[,c("Demo_Age","Demo_Male","Demo_White","Demo_SES","Demo_Married","HADS_A","HADS_D","Sleep_hours","Exercise_mins","BFI_O","BFI_C","BFI_E","BFI_A","BFI_N","Screentime_mins","SocialSupport","DERS")])

dataset1_descriptives1 = dataset1_descriptives %>%
  select("n", "mean", "sd", "min", "max","skew","kurtosis") #can be changed according to what descriptives are needed.

dataset1_descriptives1
```

# Cronbach's Alpha
```{r}
cronalpha <- alpha(select(dataset1,GAD_2,GAD_3,GAD_4,GAD_5))
```


# Correlation
For correlation, we usually create a correlation table using the corrtab package!
```{r}
corrtab <- correlation_matrix(dataset1, #the dataset
                              type="pearson", #can be pearson or spearman
                              digits=3, #number of decimal places
                              decimal.mark=".", 
                              use="lower", 
                              show_significance=TRUE,
                              replace_diagonal=FALSE,
                              replacement="" )

#write.csv(corrtab,"correlationtable.csv") #<- to export corrtab to excel format

#displaying correlation table in themed rmd!
kable(corrtab, format = "html", digits = 3) %>%
  column_spec (2:18, border_left = T, border_right = T) %>%
    row_spec(0, background = "#1D1F20") %>%
  kable_styling(latex_options = c("hold_position","scale_down")) %>%
  scroll_box(width = "800px", height = "100%")

```

# Means Tests
```{r}
ttest <- t.test(Demo_Age ~ Demo_Male, # 
                data = dataset1,
                paired = FALSE, #TRUE = dependent samples t test
                var.equal = TRUE #FALSE = Welch's t test
                )

ttestES <- effectsize::effectsize(ttest)
```

# Chi Squared Tests
```{r}
chisq <- chisq.test(Demo_Age ~ Demo_Male, # 
                data = dataset1,
                paired = FALSE, #TRUE = dependent samples t test
                var.equal = TRUE #FALSE = Welch's t test
                )
chisqES <- effectsize::effectsize(ttest)
```

# General Linear Models


## Linear Regression
